Drought indices of water demand and supply, soil moisture, vegetation, and its impact on LULCC in continental Chile

### 2024-01-15 article/paper_drought_Chile2023.tex
avanzar en paper metodologia

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  number,
  preprint,
  3p]{elsarticle}

@@

\usepackage{lineno}\linenumbers
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\journal{Journal Name}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-num}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Assessment of drought impact over land cover in continental Chile by the analysis of water supply and demand from ERA5-Land and MODIS},
  pdfauthor={Francisco Zambrano},
  pdfkeywords={drought, land cover change, satellite},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{frontmatter}
\title{Assessment of drought impact over land cover in continental Chile
by the analysis of water supply and demand from ERA5-Land and MODIS}
\author[1]{Francisco Zambrano%
\corref{cor1}%
\fnref{fn1}}
 \ead{francisco.zambrano@umayor.cl}

@@

\begin{keyword}
    drought \sep land cover change \sep 
    satellite
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, interior hidden, enhanced, breakable, sharp corners, frame hidden]}{\end{tcolorbox}}\fi

@@

Continetal Chile has a diverse climate condition from north to south and
east to west \citep{Aceituno2021} (Figure~\ref{fig-studyArea}), which
determines its great ecosystem diversity (Figure~\ref{fig-LCprop}). The
Andes Mountains are a main factor in latitudinal variation
\citep{Garreaud2009}. To describe the climate and ecosystem of Chile, we
use the Koppen-Geiger release by \citep{Beck2023} and the landcover type
persistance of 80\% of times for 2001--2022, from the IGBP
classification scheme \citep{Friedl2019}. ``Norte Grande'' and ``Norte
Chico'' predominate in an arid desert climate with hot (Bwh) and cold
(Bwk) temperatures. At the south of ``Norte Chico,'' the climate changes
to an arid steppe with cold temperatures (Bsk). Mainly, the land is
barren, with a minor surface of vegetation types such as shrubland and
grassland. In the zones ``Centro'' and the north half of ``Sur,'' the
main climate is Mediterranean, with warmer to hot summers (Csa and Csb).
In ``Centro,'' there is a major proportion (\textasciitilde50\%) of
chilean matorral (shrubland and savanna), followed by grassland
(\textasciitilde16\%), forest (8\%), and croplands (5\%). The south part
of ``Centro'' and the north part of ``Austral'' are dominated by an
Oceanic climate (Cfb). Those zones are high in forest and grassland. The
southern part of the country has a tundra climate, and in Patagonia, it
is a cold semi-arid area with an extended surface of grassland.

@@

For water supply and demand variabes, we used ERA5-Land
\citep{MunozSabater2021}, a reanalys dataset wich provides the evolution
of land variables since 1950. It has a spatial resolution of 0.1°,
hourly frequency and global coverage. We selected the variables for
total precipitation, 2 meters temperature, and volumetric soil water
layers between 0 and 100cm of depth (layer 1 to layer 3).

@@

@@

\hypertarget{impact-fof-water-supply-and-demand-and-soil-moisture-in-vegetation-productivity-within-landcover-types}{%
\subsection{Impact fof water supply and demand, and soil moisture in
vegetation productivity within landcover
types}\label{impact-fof-water-supply-and-demand-and-soil-moisture-in-vegetation-productivity-within-landcover-types}}

@@

\caption{\label{fig-LCprop}Proportion of land cover class from the
persistent land cover for 2001-2021 (\textgreater80\%) per macrozone}

@@

\hypertarget{total-water-storage-and-drought-indices}{%
\subsection{Total water storage and drought
indices}\label{total-water-storage-and-drought-indices}}

Terrestrial Water Storage (TWS) is defined as the total amount of water
stored on land that includes any natural or artificial water bodies,
such as groundwater, soil moisture, rivers, lakes, snowpack, ice, and
biomass water \citep{Humphrey2023, Deng2023}. TWS changes evidence the
effects of multiple water fluxes on the hydrological cycle
\citep{Deng2023}. These are reflected in the temporal variation of
observations of the Earth's gravitational field
\citep{Abolafia2021, Sabzehee2023} and, in recent decades, the twin
Gravity Recovery and Climate Experiment (GRACE) satellites and their
Follow-On mission (GRACE-FO) have provided valuable results on globally
distributed TWS anomalies \citep{Tapley2019, Ferreira2023}. The GRACE
mission was launched in March 2002 and was operational until October
2017 \citep{Ramjeawon2022}. Then its GRACE Follow-On successor was
launched in May 2018 \citep{Landerer2020, Yin2022}. The information
provided by these satellites is used to construct monthly maps of the
Earth's average gravity field, providing details of the movement of
water masses or water mass anomaly estimates relative to the long-term
average gravity field \citep{Humphrey2023, Wahr2004}.

In this study, RL06.1\_V3 GRACE mascon (mass concentration) monthly
solutions were used, which are provided by the Jet Propulsion Laboratory
(JPL-M, https://grace.jpl.nasa.gov). Each GRACE Tellus monthly grid at
0.5 degrees represents the deviation of surface mass for that month
relative to a reference time average (2004-2009 baseline), which is
subtracted from all other monthly grids to provide terrestrial water
storage anomalies (TWSA) \citep{Ramjeawon2022, Yin2022}. This JPL-M
version of the data employs a Coastal Resolution Improvement (CRI)
filter that reduces signal leakage errors across coastlines
\citep{NASA2019, Wiese2016}. Although the mascon solutions greatly
reduced leakage errors, a gain factor, which is used to enhance the
spatial resolution, was applied to the dataset as recommended for
hydrological studies \citep{Ramjeawon2022, Yin2022}. The water storage
and height anomalies are given in Equivalent Water Height or Thickness
units (EWH, cm) \citep{Sabzehee2023}, and their temporal resolution
monthly is from April 2002 to May 2023.

@@

\hypertarget{trend-of-drought-indices-for-water-demandsupply-soil-moisture-and-vegetation-productivity-1}{%
\subsection{Trend of drought indices for water demand/supply, soil
moisture, and vegetation
productivity}\label{trend-of-drought-indices-for-water-demandsupply-soil-moisture-and-vegetation-productivity-1}}

@@

\hypertarget{lulc-change-for-2001-2022-and-its-relation-with-water-supply-and-demand-and-soil-moisture-1}{%
\subsection{LULC change for 2001-2022 and its relation with water supply
and demand, and soil
moisture}\label{lulc-change-for-2001-2022-and-its-relation-with-water-supply-and-demand-and-soil-moisture-1}}

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  number,
  preprint,
  3p,
  onecolumn]{elsarticle}

@@

\usepackage{lineno}\linenumbers \usepackage{multirow}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\journal{Journal Name}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-num}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Drought indices of water demand and supply, soil moisture, vegetation, and its impact on LULCC in continental Chile},
  pdfauthor={Francisco Zambrano},
  pdfkeywords={drought, land cover change, satellite},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{frontmatter}
\title{Drought indices of water demand and supply, soil moisture,
vegetation, and its impact on LULCC in continental Chile}
\author[1]{Francisco Zambrano%
\corref{cor1}%
\fnref{fn1}}
 \ead{francisco.zambrano@umayor.cl}

@@

\begin{keyword}
    drought \sep land cover change \sep 
    satellite
\end{keyword}
\end{frontmatter}
    \captionsetup{justification=raggedright,singlelinecheck=false}

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, frame hidden, boxrule=0pt, breakable, borderline west={3pt}{0pt}{shadecolor}, sharp corners, interior hidden]}{\end{tcolorbox}}\fi

@@

Continetal Chile has a diverse climate condition from north to south and
east to west \citep{Aceituno2021} (Figure~\ref{fig-studyArea}), which
determines its great ecosystem diversity (Figure~\ref{fig-LCprop}). The
Andes Mountains are a main factor in latitudinal variation
\citep{Garreaud2009}. To describe the climate and ecosystem of Chile, we
use the Koppen-Geiger release by \citep{Beck2023} and the landcover type
persistance of 80\% of times for 2001--2022, from the IGBP
classification scheme \citep{Friedl2019} (see
Section~\ref{sec-methods_lulc}). ``Norte Grande'' and ``Norte Chico''
predominate in an arid desert climate with hot (Bwh) and cold (Bwk)
temperatures. At the south of ``Norte Chico,'' the climate changes to an
arid steppe with cold temperatures (Bsk). Mainly, the land is barren,
with a minor surface of vegetation types such as shrubland and
grassland. In the zones ``Centro'' and the north half of ``Sur,'' the
main climate is Mediterranean, with warmer to hot summers (Csa and Csb).
In ``Centro,'' there is a major proportion (\textasciitilde50\%) of
chilean matorral (shrubland and savanna), followed by grassland
(\textasciitilde16\%), forest (8\%), and croplands (5\%). The south part
of ``Centro'' and the north part of ``Austral'' are dominated by an
Oceanic climate (Cfb). Those zones are high in forest and grassland. The
southern part of the country has a tundra climate, and in Patagonia, it
is a cold semi-arid area with an extended surface of grassland.

@@

For water supply and demand variabes, we used ERA5-Land
\citep{MunozSabater2021}, a reanalys dataset wich provides the evolution
of land variables since 1950. It has a spatial resolution of 0.1°,
hourly frequency and global coverage. We selected the variables for
total precipitation, 2 meters temperature maximum and minimum, and
volumetric soil water layers between 0 and 100cm of depth (layer 1 to
layer 3).

@@

\begin{table}[!ht]
\caption{Description of the earth observation data used }
\label{tab-desEOD}
\small
\centering
\begin{tabular}{p{0.13\textwidth}cp{0.3\textwidth}p{0.095\textwidth}ccc}
\hline
\multirow{1}{*}{\centering Product} & Sub-product & Variable & Spatial Resolution  & Period & Units & Short Name \\ 
\hline
\multirow{4}{*}{ERA5-Land} & ~ & Precipitation & \multirow{4}{*}{~0.1°} & \multirow{4}{*}{1981-2023} & mm & P \\ 
         &  & Maximum temperature & ~ & & $°C$ & $T_{max}$ \\ 
         &  & Minimum temperature & ~ & & $°C$ & $T_{min}$ \\ 
         &  & Volumetric Soil Water Content at 1m & ~ & & $m3/m3$ & SM \\ 
ERA5-Land* & & Atmospheric Evaporative Demand & 0.1° & 1981-2023 & mm & AED \\
        \multirow{2}{*}{MODIS} & MOD13A3.061 & Normalized Difference Vegetation Index & \multirow{2}{*}{~1 km} & 2000-2023 & ~ & NDVI \\ 
         & MCD12Q1.061 & Landcover IGBP scheme & & 2001-2022 & ~ & Landcover \\ 
\hline
\end{tabular}
{\raggedright *Derived from ERA5-Land with Eq. \ref{eq-AED}. \par}
\end{table}

@@

\hypertarget{software-and-packages-used}{%
\subsection{Software and packages
used}\label{software-and-packages-used}}

For the downloading, processing, and analysis of the spatio-temporal
data, we used the open source software for statistical computing and
graphics, \texttt{R} \citep{R2023}. For downloading ERA5-Land, we used
the \texttt{\{ecmwfr\}} package \citep{Hufkens2019}. For processing
raster data, we used \texttt{\{terra\}} \citep{Hijmans2023} and
\texttt{\{stars\}} \citep{Pebesma2023}. For managing vectorial data, we
used \texttt{\{sf\}} \citep{Pebesma2018}. For the calculation of AED, we
used \texttt{\{SPEI\}} \citep{Bergueria2023}.

\hypertarget{drought-indices}{%
\subsection{Drought Indices}\label{drought-indices}}

We derived the drought indices of water supply and demand, soil moisture
from the ERA5-Land dataset, and vegetation from the MODIS product, all
at monthly frequency.

For the indices EDDI and SPEI that use water demand, first we have to
calculate the AED. For this, we used the method of Hargreaves
\citep{Hargreaves1994}:

\begin{equation}\protect\hypertarget{eq-AED}{}{AED = 0.0023\cdot Ra\cdot (T+17.8)\cdot (T_{max}-T_{min})^{0.5}}\label{eq-AED}\end{equation}

where \(Ra\) \((MJ\,m^2\, day^{-1})\) is extraterrestrial radiation;
~\(T\), \(T_{max}\), and \(T_{min}\) are mean, maximum, and minimum
temperature \((°C)\). We calculate the centroid coordinates per pixel
and use the latitude to estimate \(Ra\).

To evaluate water demand, we chose the \(EDDI\)
\citep{Hobbins2016, McEvoy2016} index, which uses the \(AED\). For
supply, we used the index recommended by the World Meteorological
Organization (WMO) for monitoring drought, the SPI \citep{Mckee1993}. We
calculated the SPEI, which used a balance between \(P\) and \(AED\), in
this case, an auxiliary variable \(D = P-AED\) is used. In this study,
we propose the \(zcSM\) (standardized anomaly of acumulated soil
moisture at 1 m), which uses SM. Finally, for the proxy of productivity,
\(zcNDVI\), we used the NDVI. Before using the NDVI, it was smoothed
following the procedure described in \citep{Zambrano2018} and
\citep{Zambrano2016}.

All the indices are multi-scalar and were calculated for time scales of
1, 3, 6, 12, 24, and 36 months, except for zcNDVI, which was calculated
for 6 months. The goal is to be able to evaluate short- and long-term
droughts in water demand and supply and soil moisture. This is
particularly important for central Chile because it has suffered from a
prolonged decrease in precipitation for more than 12 years
\citep{Garreaud2020, Boisier2018, Garreaud2017}.

To calculate the drought indices, first we must calculate the
accumulation of the variable. In this case, for generalization purposes,
we will use \(V\), referring to \(P\), \(AED\), \(D\), \(NDVI\), and
\(SM\) (Table \ref{tab-desEOD}). We cumulated each \(V\) over the time
series of \(n\) values, and for the time scales \(s\):

\begin{equation}\protect\hypertarget{eq-sumvar}{}{A_{si} = \sum_{i=n-s-i+2}^{n-i+1} V_i\,\, \forall\, i\geq n-s+1  }\label{eq-sumvar}\end{equation}

It corresponds to a moving window (convolution) that sums the variable
for \(s\) starting for the last month \(n\) until the month, which could
sum for \(s\) months (n-s+1). Once the variable is cumulated over time
for the scale \(s\), we used a nonparametric approach following
\citep{Hobbins2016} to derive the drought indices. Thus, the empirically
derived probabilities are obtained through an inverse normal
approximation \citep{Abramowitz1968}. Then, we used the empirical Tukey
plotting position \citep{Wilks2011} over \(A_i\) to derive the
\(P(A_i)\) probabilities across a period of interest:

\begin{equation}\protect\hypertarget{eq-probPai}{}{P(A_i) = \frac{i-0.33}{n+0.33'}}\label{eq-probPai}\end{equation}

The drought indices \(SPI\), \(SPEI\), \(EDDI\), \(zcSM\), and
\(zcNDVI\) are obtained following the inverse normal approximation:

\begin{equation}\protect\hypertarget{eq-DI}{}{DI(A_i) = W - \frac{C_0+C_1\cdot W + c_2 \cdot W^2}{1+d_1\cdot W +d_2\cdot W^2 +d_3\cdot W^3}}\label{eq-DI}\end{equation}

\(DI\) is referring to the drought index calculated for the variable
\(V\). The values for the constats are: \(C_0 = 2.515517\),
\(C_1 = 0.802853\), \(C_2 = 0.010328\), \(d_1 = 1.432788\),
\(d_2 = 0.189269\), and \(d3 = 0.001308\). For \(P(A) \leq 0.5\),
W=\(\sqrt{-2\cdot ln(P(A_i))}\) , and for \(P(A_i) > 0.5\), replace
\(P(A_i)\) with \(1-P(A_i)\) and reverse the sign of \(DI(A_i)\).

\hypertarget{trend-of-drought-indices-for-water-demand-and-supply-soil-moisture-and-vegetation-productivity}{%
\subsection{Trend of drought indices for water demand and supply, soil
moisture, and vegetation
productivity}\label{trend-of-drought-indices-for-water-demand-and-supply-soil-moisture-and-vegetation-productivity}}

To estimate if there are significant positive or negative trends for the
drought indices, we used the non-parametric test of Mann-Kendall
\citep{Kendall1975}. To determine the magnitude of the trend, we used
Sen's slope \citep{Sen1968}. One of the advantages of applying this
methodology is that the Sen's slope is not affected by outliers as
regular regression does. We applied both to the five indices at the six
time scales from 1981 to 2023 and indices SPI, EDDI, SPEI, and zcSM, and
from 2000 to 2023 in the case of zcNDVI.

\citep{Vicente-Serrano2021} made a global analysis of the drought's
severity trend using SPI, SPEI, and the Standardized Evapotranspiration
Deficit Index (SEDI; \citep{Vicente-Serrano2018}) to evaluate AED. They
indicate that the increase in hydrological drought has been due to
anthropogenic effects rather than climate change. This is because the
global increase in AED does not explain the change in the spatial
pattern of hydrological drought. Also, they state that \emph{``the
increase in hydrological droughts has been primarily observed in regions
with high water demand and land cover change''}. We will contrast this
hypothesis with what is occurring in Chile. To achieve this, we will use
the landcover class type that remain more than 80\% of types for
2001-2022, and use this as a mask, where there are low changes, to
evaluate the trend on zcNDVI.

\hypertarget{sec-methods_lulc}{%
\subsection{LULC change for 2001-2022 and its relation with water supply
and demand, and soil moisture}\label{sec-methods_lulc}}

@@

\caption{\label{fig-LCprop}Proportion of land cover class from the
persistent land cover for 2001-2022 (\textgreater80\%) per macrozone}

@@

To analyze the LULCC, we use the IGBP scheme from the MCD12Q1 collection
6.1 from MODIS. This product has a yearly frequency from 2011 to 2022.
The IGBP defines 17 classes; from these, we regrouped into ten
macroclasses for forest, schrublands, savannas, grasslands, wetlands,
croplands, urban, snow and ice, barren, and water bodies. Then, we
calculated the surface ocuppied per landcover class into each macrozone
(``Norte Grande'' to ``Austral'') per year. To determine the trend of
change we used the Sen' slope \citep{Sen1968} over the surface per year.

We calculated a raster mask considering the classes that remains more
than 80\% of years (2001-2022), which allow us to identify the areas
with no landcover change for the macroclasses. Figure~\ref{fig-LCprop}
shows the summary of proportion of surface per landcover class and
macrozone, derived from the mask over continental Chile.

We will explore if the trend in landcover classess is associated to
trend of the drought indices. For this we will use the regularizations
techniques of Lasso \citep{Tibshirani2010} and Ridge regressión
\citep{Hoerl1970}. Also, we will test random forest for this purpose
\citep{Ho1995}.

\hypertarget{impact-for-water-supply-and-demand-and-soil-moisture-in-vegetation-productivity-within-landcover-types}{%
\subsection{Impact for water supply and demand, and soil moisture in
vegetation productivity within landcover
types}\label{impact-for-water-supply-and-demand-and-soil-moisture-in-vegetation-productivity-within-landcover-types}}

The aim of this section is to analyze the drought indices of water
demand and supply and soil moisture against vegetation to address: i) if
short- or long-term time scales are most important in impacting
vegetation through Chile; and ii) the strength of the correlation for
the variable and the time scale. Then, we will summarize for each
landcover class and macrozone. Thus, we will be able to advance in
understanding how climate is affecting vegetation, considering the
impact on the five macroclasses: forest, cropland, grassland, savanna,
and shrubland.

To assess how water demand and supply and soil moisture are related to
vegetation productivity (zcNDVI), we analyze the linear correlation
between the indices SPI, SPEI, EDDI, and zcSM for 1, 3, 6, 12, 24, and
36-month time scales against zcNDVI. We followed a similar approach to
that used by \citep{Meroni2016} when using the SPI for meteorological
drought against the cumulative FAPAR (Fraction of Absorbed
Photosynthetically Active Radiation) as a proxy for vegetation
productivity. We made a pixel-to-pixel linear correlation analysis per
index. First, we calculate the Pearson coefficient of correlation for
the six time scales and let the time scale that reaches the maximum
correlation be significant (p \textless{} 0.05). Then, we extracted the
Pearson correlation value corresponding to the time scales that reached
the maximum value. Thus, we derived two raster maps per index, the first
with the time scales and the second with the correlation value.

@@

@@
